\LARGE \bf Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects

### 2022-11-24 PaperPreliminary.tex
.

% System overview
    \section{System overview}
    Considering the goal of evaluating reference spreading in a practial usecase, this work will focus on a dual-arm robotic setup. Having two arms increases the maximum payload. Furthermore, some object manipulation tasks, such as grabbing, require contact from multiple sides. The impact tasks that are considerd in this work are stamping, swiping, grabbing, and tilting.

@@

    \subsection{Robot dynamics}
    The robot dynamics can be modeled as 
    \begin{equation}
    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^TF_{c,i}
    \end{equation}

    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$. The robot has $n$ contacts, and for each contact i there is an external contact wrench $F_{c,i}$. The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocity of the contact body and $\dot{p}$ the carthesian velocity of the contact point.

    For control purposes, it is assumed that contact forces only act on the end effector body. These forces are modeled as if they act at the same control point $p$, which is chosen as the intersection of robot link 5 and 7. Furthermore, omitting dependency on $q$ and $\dot{q}$ for brevity results in

    \begin{equation}
    M\ddot{q}+h=\tau_{cmd}+J^TF_c.
    \end{equation}

    \subsection{Base controller}
    For articulated robot arms, it is convenient to control the pose of the end effector, rather than controlling the joint angles. This can be accomplised with an impedance controller.

    The desired impedance behaviour is
    \begin{equation}
     \Lambda \begin{bmatrix} \dot{\omega} - \dot{\omega}_{t} \\ \ddot{p} - \ddot{p}_{t} \end{bmatrix} + D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} = F_c-F_{c,t}:
     \end{equation}

    a mass spring damper with with task-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx. Subscript $t$ denotes a target.

    Stifness $K$ is typically chosen as a diagonal matrix. The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric. Furthermore, For the inertia matrix, two options were considered. Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t. to $e$, resulting in better tracking in free motion. In this work however, the task-space inertia is set to match the joint-space inertia following $\Lambda = JM^{-1}J$, decoupling the contact force w.r.t. $e$ for better performance during contact. Further motivation for this decision is provided in Appendix xxx. % = \begin{bmatrix} e_{rotation} & e_{translation} \end{bmatrix}^T $.

    Based on (x), the desired task-space accelerations are given by

    \begin{equation}
    \begin{bmatrix} \dot{\omega}\\ \ddot{p}\end{bmatrix} = \begin{bmatrix} \dot{\omega}_{t}\\ \ddot{p}_{t}\end{bmatrix} + \Lambda^{-1} \left (  D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} +F_{c,t} \right ).
    \end{equation}

    Note the exclusion of $F_c$ as the exerted contact wrench is not modeled and therefore unknown. Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics and safety constraints such as maximum joint torques and velocities. The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot. To resolve this redundancy, one more task which describes the desired acceleration of the first robot joint is added. This so-called posture task with stiffness $k_p$ is given by

    \begin{equation}
    \ddot{q}_1 = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,t}).
    \end{equation}
    An overview of the used control parameters is given in Table xxx.

    \begin{table}[h]
    \centering
    \begin{tabular}{l|l}
    \hline
    Parameter & Value                                                \\ \hline
    $q_{1,t}$     & 0 rad                                              \\
    $k_1$     & 50 Nm/rad                                            \\
    $k_2$     & 800 N/m                                              \\
    $k_p$     & 50 Nm/rad                                            \\
    $K$       & $\begin{bmatrix} k_1 I &0 \\ 0 & k_2 I\end{bmatrix}$
    \end{tabular}
    \end{table}

@@

% Teleoperation
    \section{Teleoperation}

>>>>

% System overview
    \section{System overview}
    Considering the goal of evaluating reference spreading in a practical usecase, this work will focus on a dual-arm robotic setup. Having two arms increases the maximum payload. Furthermore, some object manipulation tasks, such as grabbing, require contact from multiple sides. The impact tasks that are considered in this work are stamping, swiping, grabbing, and tilting.

@@

    \subsection{Robot dynamics}
    The robot dynamics can be described as 
    \begin{equation}
    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^Tf_{c,i}
    \end{equation}

    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$. The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$. The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $\dot{p}$ the Cartesian velocities of the contact point.

    For modeling purposes, it is assumed that contact forces only act on the end effector body. These forces are modeled as a single contact wrench $f_c$ acting on control point $p$, which is chosen as the intersection of robot link 5 and 7. This assumption allows for $f_c$ to be estimated using methods described in section xxx. Furthermore, omitting dependency on $q$ and $\dot{q}$ from for brevity results in

    \begin{equation}
    M\ddot{q}+h=\tau_{cmd}+J^Tf_c.
    \end{equation}

    \subsection{Base controller}
    For articulated robot arms, it is convenient to control the pose of the end effector in Cartesian space, rather than controlling the joint angles. This can be accomplished with an impedance controller.

    The desired impedance behavior is
    \begin{equation} \label{eq:impedance_desired}
     \Lambda  \begin{bmatrix} d_{\dot{\omega}} - \dot{\omega}   \\ d_{\ddot{p}} - \ddot{p}  \end{bmatrix} + D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix} = d_{f_c} -f_c:
     \end{equation}

    a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx. Desired values are denotes as $d_{(\cdot)}$, e.g., the desired value for $\ddot{p}$ is  $d_{\ddot{p}}$.

    Stiffness $K$ is typically chosen as a diagonal matrix. The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric. Furthermore, For the inertia matrix, two options were considered. Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t. to the position error, resulting in better tracking in free motion. This approach was used in~\cite{vanoorschotDesignNumericalValidation2022}. In this work, however, the task-space inertia is set to match the joint-space inertia following $\Lambda^{-1} = JM^{-1}J^T$, decoupling the contact force w.r.t. the position error for better performance during contact. Further motivation for this decision is provided in Appendix xxx. % = \begin{bmatrix} e_{rotation} & e_{translation} \end{bmatrix}^T $.

    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $t_{\dot{\omega}}$ and $t_{\ddot{p}}$, following

    \begin{align} \begin{bmatrix}t_{\dot{\omega}}\\ t_{\ddot{p}}\end{bmatrix} &= \begin{bmatrix} {\dot{\omega}}\\{\ddot{p}}\end{bmatrix}+\Lambda^{-1} f_c\\
     &=     \begin{bmatrix} d_{\dot{\omega}}    \\ d_{\ddot{p}}  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix}- d_{f_c} \right ).
    \end{align}

    Note the exclusion of $f_c$ as the exerted contact wrench is not modeled and therefore unknown. Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics (x) and safety constraints (x).

    The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot. To resolve this redundancy, one more task which describes the target acceleration of the first robot joint is added. This so-called posture task with stiffness $k_p$ is given by

    \begin{equation}
    \ddot{q}_{1,t} = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,d}).
    \end{equation}
    An overview of the used control parameters is given in Table xxx.

    \begin{table}[h]
    \centering
    \begin{tabular}{l|l}
    \hline
    Parameter & Value                                                \\ \hline
    $q_{1,d}$     & 0 rad                                              \\
    $k_1$     & 50 Nm/rad                                            \\
    $k_2$     & 800 N/m                                              \\
    $k_p$     & 50 Nm/rad                                            \\
    $K$       & $\begin{bmatrix} k_1 I &0 \\ 0 & k_2 I\end{bmatrix}$
    \end{tabular}
    \end{table}

@@

% Teleoperation
    \section{Trajectory planning with teleoperation}

### 2022-11-24 Progress_12.tex
.

% Summary 
  \section*{Progress meeting 11 | 4 November, 2022}

@@

  \begin{itemize}
      \item Reference spreading controllers:
        Last time I was having issues with getting the controller to work with a force reference. This was a programming issue which has since been fixed. In II.B of the paper attached at the end of this report, I give the general description of this controller. In IV.B, the control behaviour during ante, intermediate, and post impact phase are further specified. Section V shows the results. Here is are some things I found interesting about the results:
        \begin{itemize}
          \item Reference spreading reduces the control effort during non-nominal impacts
          \item Without a way to detect when contact is completed, introducing an intermediate impact phase with a predetermined length is not always beneficial.
          \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
        \end{itemize}
      \item Writing: So far I have written part of the introduction, an explanation of the impedance QP controller, and various options for the intermediate impact phase controller. The draft paper is attached at the end of this document.

@@

  \item[2b] \textbf{Contact monitoring:} When investigating contact monitoring, two approaches can be taken: either using proprioceptive or exteroceptive sensors. A possible improvement for contact monitoring using proprioceptive sensors could be to wait a fixed time starting from the last detected impact, rather than waiting a fixed time from the first impact. As for exteroceptive sensors, they can be a hurdle for large-scale commercial applications as they are not integrated in the robot. However, if a soft cover is to be mounted to the end effector, including tactile sensors for impact and contact monitoring becomes more feasible. Practical questions such as which tactile sensor to use and how to integrate it could be addressed, though this is not absolutely necessary for completing the research goals, and therefore has a low priority.
  \end{enumerate}
% main text
\newpage
  \section{Reference spreading results}
  The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix). This setup was evaluated during a stamping task on the panda setup. Figure \ref{fig:my_label} shows the results.

>>>>

% Summary 
  \section*{Progress meeting 12 | 18 November, 2022}

@@

  \begin{itemize}
      \item Reference spreading controllers:
        Last time I was having issues with getting the controller to work with a force reference. This was a programming issue which has since been fixed. In II.B of the paper attached at the end of this report, I give the general description of this controller. In IV.C, the control behaviour during ante, intermediate, and post impact phase are further specified. Section \ref{sec:results} shows the results. Here is are some things I found interesting about the results:
        \begin{itemize}
          \item Reference spreading reduces the control effort during non-nominal impacts
          \item Without a way to detect when contact is completed, introducing an intermediate impact phase with a predetermined length is not always beneficial.
          \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
          \item A newly proposed intermediate control approach effectively blends the ante- and post impact reference for a balance between tracking and control effort.
        \end{itemize}
      \item Writing: So far I have written part of the introduction, an explanation of the impedance QP controller, and various options for the intermediate impact phase controller. The draft paper is attached at the end of this document.

@@

  \item[2b] \textbf{Contact monitoring:} When investigating contact monitoring, two approaches can be taken: either using proprioceptive or exteroceptive sensors. A possible improvement for contact monitoring using proprioceptive sensors could be to wait a fixed time starting from the last detected impact, rather than waiting a fixed time from the first impact. As for exteroceptive sensors, they can be a hurdle for large-scale commercial applications as they are not integrated in the robot. However, if a soft cover is to be mounted to the end effector, including tactile sensors for impact and contact monitoring becomes more feasible. Practical questions such as which tactile sensor to use and how to integrate it could be addressed, though this is not absolutely necessary for completing the research goals, and therefore has a low priority.
  \end{enumerate}
% main text
\newpage
  \section{Reference spreading results}\label{sec:results}
  The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix). This setup was evaluated during a stamping task on the panda setup. Figure \ref{fig:stamp} shows the results.

  The command force $F_{cmd}$ will be used to evaluate the control effort. Without reference spreading, there is a clear peak in controll effort related to the peak in the velocity tracking error. The option RS inter0 does not suffer from this peak due to switching of the reference. None of the intermediate modes outperform RS inter0 for two reasons. Firstly, there is only one clearly visible impact while the intermediate mode is supposed to address near-simultaneous impacts. Furthermore, it is assumed that the intermediate mode lasts 0.2 seconds, whereas contact is completed sooner than that.

  Similarly, the controller was tested for a grabbing task using two arms. The results for the individual arms is shown in Figure \ref{fig:grab1} and \ref{fig:grab2}. In this scenario, the effect of the intermediate mode is more obvious since there are near-simultaneous impacts: the impact of arm 1, and the impact of arm 2. Intermediate options 1 and 2 experience a peak in the position tracking error which is based on the post-impact reference. This problem is resolved in option 3 thanks to the blending. Intermediate option 3 arguably has the best tradeoff between tracking performance and controll effort moderation.

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of a stamping experiment. The top figure shows the results for the various approaches, both over the entire experiment, and over a small timeframe around the impact. The bottom figure isolates each approach for additional clarity. The dotted lines indicate the target as described in the control approaches.}
  \label{fig:stamp}
  \end{figure}

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of arm 1 in boxgrab experiment. (no RS and RS inter0 show similar results. This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing. This requires further investigation.)}
  \label{fig:grab1}
  \end{figure}

\begin{figure}[]
  \centering
  \begin{adjustwidth}{0pt}{5pt}
  \centering
  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_1.pdf}\\
    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_2.pdf}
  \end{adjustwidth}
  \caption{Results of arm 2 in boxgrab experiment.(no RS and RS inter0 show similar results. This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing. This requires further investigation.)}
  \label{fig:grab2}
  \end{figure}
